---
title: "Modeling"
author: "Marton Barta & David Torgerson"
date: "4/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Set Up - Calling the Libraries
```{r}
library(Amelia)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(reshape2)
library(naniar) # Load nanair for missing data visualization
library(OneR) # Load OneR for binning function
library(fastDummies) # for dummy variable creation
library(lubridate) # for date to month conversion
library(glmnet)
library(plotmo)
library(rpart)
library(caret) # for confusion matrix
library(randomForest)
library(xgboost)
library(pROC) # Load proc
library(ggforce)
```

## Importing the Data
```{r Read Data Set}
data <- read.csv("/Volumes/GoogleDrive/Shared drives/HPA Spring 2021/data.csv")
```

## Data Preparation

First lets work through some data preparation. As a first step we can view the dataset and summarize it:
```{r}
head(data)
tail(data)
colnames(data)
summary(data)
dim(data)
```

We have 6012 samples in the data and 144 variables. The dependent or response variable is: Winner.

## Response Variable:
```{r}
## Response Variable Values
unique(data$Winner)

## Check Response Variable Frequency
as.data.frame(table(data$Winner))
```
There are 110 matches when the result ended up being a draw. For the simplicity of our project and considering the fact that draws are very unlikely to occur, we will get rid of those matches completely for the rest of the project.

The frequency seems to be a little unevenly distributed, but we have enough Blue Corner winners to continue our project.

```{r}
## Getting rid of the draw matches
data<-subset(data, Winner!="Draw")
as.data.frame(table(data$Winner))
```
Descriptions of our response values: 
Blue – The Blue corner fighter won the match.
Red - The Red corner fighter won the match.


We will also get rid of the Women matches:
```{r}
## Getting rid of the Women weight classes
data<-subset(data, weight_class!="WomenStrawweight")
data<-subset(data, weight_class!="WomenBantamweight")
data<-subset(data, weight_class!="WomenFlyweight")
data<-subset(data, weight_class!="WomenFeatherweight")
as.data.frame(table(data$weight_class)) # checking the rest of the weight classes
```

## Checking Missing values
```{r}
missmap(data, col=c("white", "black"), legend=FALSE)
```
The graph indicates some sort of relationship between missing data. However, most of the data seems to be present, which is promising for our analysis.

Next, we will examine where our missing values are, and try to figure out the reason behind them.
```{r}
sum(is.na(data))
colSums(is.na(data))
```
We have a total of 95,614 instances of missing data, but it is fairly well distributed - we have 1275 missing values for 49(!) variables. This means that 65% of the missing values are accounted for by those variables, while the rest can be explained by the ones that are missing 646 values and some other missing values across our dataset.

```{r}
str(data)
```
As we can see, we have all kinds of datatypes (chr, num, int); therefore, we have to be careful when we are working in the followings. We will have to be mindful about this since we will not be able to work with quantitative models just now.

## Organizing the variable names and structuring the dataset

```{r}
colnames(data)
```
Our dataset has an unorganized structure; working with these variables would be hard, so, we will change the variable names. As we see, the dataset contains features to the Red (R_) and Blue (B_) fighters. We want to separate the fighters, so we can analyze them not just as opponents, but also as our picked fighter. In the following, we will fix the variable names and we will separate the fighters, meaning we will end up with double sample size.

Wherever the variable starts with 'B_' or 'R_' (these refer to the fighters), we will get rid of those characters and we will organize those features into a variable which is associated with either the red or blue fighter.
```{r}
## Blue Fighter
b_cols <- names(data)[grep("B_",names(data))][c(1:68,71)]
#b_cols
b_dat <- data[,b_cols]
#b_dat
names(b_dat) <- str_replace(names(b_dat),"B_", "")
#names(b_dat)

## Red Fighter
r_cols <- names(data)[grep("R_",names(data))][c(1,12:79)]
#r_cols
r_dat <- data[,r_cols]
#r_dat
names(r_dat) <- str_replace(names(r_dat),"R_", "")
#names(r_dat)
```

Next, we will add the corresponding opponent data to both the Red and Blue datasets so that each row has the match data for both the fighter and their opponent. This will be helpful for training our model later on.
```{r}
opp_1 <- r_dat
#opp_1
names(opp_1) <- paste(names(opp_1), "_opp")
opp_2 <- b_dat
names(opp_2) <- paste(names(opp_2), "_opp")
```

```{r}
dat_1 <- cbind.data.frame(b_dat, opp_1)
dat_1$Winner <- rep(0,nrow(dat_1))
dat_1$Winner[data$Winner == "Blue"] <- 1

dat_2 <- cbind.data.frame(r_dat, opp_2)
dat_2$Winner <- rep(0,nrow(dat_2))
dat_2$Winner[data$Winner == "Red"] <- 1
```

Now, we will combine the red and blue data into a master fighter dataset called f_data.

```{r}
f_data <- rbind.data.frame(dat_1, dat_2)
#f_data
#names(f_data)
```

Finally, we will add the admin data (such as referee, location, title_bout, and weight_class) to the fighter data into a dataset called combined. These variables did not contain neither R_ nor B_ at the beginnning of their names, meaning that we have not added those to our variable yet.

```{r}
#names(data)
admindat <-  data[,c(3,4,5,7,8)]
admindat <- rbind.data.frame(admindat,admindat)
admindat
names(admindat)
```

```{r}
combined <- cbind(f_data,admindat) 
names(combined)
```
The variable combined should contain all the variables that we will need for our analysis. This variable also lists the features as we want it to be.

These are our variables:
* `fighter` - The name of the fighter
* `_opp_` - containing columns is the average of damage done by the opponent on the fighter
* `_opp` - if it is at the end of the variable, then the variable refers to the fighter's opponent
* `__pct` - Percentage of something
* `KD` - is number of knockdowns
* `SIG_STR` - is no. of significant strikes 'landed of attempted'
* `SIG_STR_pct` - is significant strikes percentage
* `TOTAL_STR` - is total strikes 'landed of attempted'
* `TD` - is no. of takedowns
* `TD_pct` - is takedown percentages
* `SUB_ATT` - is no. of submission attempts
* `PASS` - is no. times the guard was passed?
* `REV` - is the no. of Reversals landed
* `HEAD` - is no. of significant strinks to the head 'landed of attempted'
* `BODY` - is no. of significant strikes to the body 'landed of attempted'
* `CLINCH` - is no. of significant strikes in the clinch 'landed of attempted'
* `GROUND` - is no. of significant strikes on the ground 'landed of attempted' win_by is method of win
* `last_round`- is last round of the fight (ex. if it was a KO in 1st, then this will be 1)
* `last_round_time` - is when the fight ended in the last round
* `Format` - is the format of the fight (3 rounds, 5 rounds etc.)
* `Referee` - is the name of the Ref
* `date` - is the date of the fight
* `location` - is the location in which the event took place
* `Fight_type` - is which weight class and whether it's a title bout or not
* `Winner` - is the winner of the fight
* `Stance` - is the stance of the fighter (orthodox, southpaw, etc.)
* `Height_cms` - is the height in centimeter
* `Reach_cms` - is the reach of the fighter (arm span) in centimeter
* `Weight_lbs` - is the weight of the fighter in pounds (lbs)
* `age` - is the age of the fighter
* `title_bout`  - Boolean value of whether it is title fight or not
* `weight_class` - is which weight class the fight is in (Bantamweight, heavyweight, flyweight, etc.)
* `no_of_rounds` - is the number of rounds the fight was scheduled for
* `current_lose_streak` - is the count of current concurrent losses of the fighter
* `current_win_streak` - is the count of current concurrent wins of the fighter
* `draw` - is the number of draws in the fighter's ufc career
* `wins` - is the number of wins in the fighter's ufc career
* `losses` - is the number of losses in the fighter's ufc career
* `total_rounds_fought` - is the average of total rounds fought by the fighter
* `total_time_fought(seconds)` - is the count of total time spent fighting in seconds
* `total_title_bouts` - is the total number of title bouts taken part in by the fighter
* `win_by_Decision_Majority` - is the number of wins by majority judges decision in the fighter's ufc career
* `win_by_Decision_Split` - is the number of wins by split judges decision in the fighter's ufc career
* `win_by_Decision_Unanimous` - is the number of wins by unanimous judges decision in the fighter's ufc career
* `win_by_KO/TKO` - is the number of wins by knockout in the fighter's ufc career
* `win_by_Submission` - is the number of wins by submission in the fighter's ufc career
* `win_by_TKO_Doctor_Stoppage` - is the number of wins by doctor stoppage in the fighter's ufc career

```{r}
as.data.frame(table(combined$Winner))
```
Since we have equal amount of winners and losers, that means that our transformation was successful and we ended up with a relatively big data size (10,874).

Now, we will determine the most important variables in the dataset for predicting the Winner. These are the variables with which we will train our model.
```{r Check Variable Occurence in Terms of Response Variable}
# Select variables to use
vars <- c("current_win_streak", "losses", "win_by_KO.TKO","win_by_Submission","current_lose_streak", "avg_SIG_STR_pct", "current_win_streak _opp","avg_TD_att", "current_lose_streak _opp", "longest_win_streak _opp", "wins _opp", "losses _opp", "wins")

# Create vector of responses
resps <- c("1","0")
outcome <- c("1","0")


# Create empty data frame to store proportion
res_mat <- as.data.frame(matrix(NA, nrow = length(vars), ncol = length(outcome)))

# Loop through and calculate proportion for each question and response
s_dat <- as.data.frame(scale(combined[,vars]))

names(s_dat) <- vars

for(i in 1:nrow(res_mat)){
  # Proportion of Won
  res_mat[i,1] <- mean(s_dat[combined$Winner == 1, vars[i]], na.rm = TRUE)
  # Proportion of Lost 
  res_mat[i,2] <-  mean(s_dat[combined$Winner == 0, vars[i]], na.rm = TRUE)
}


# Add column names
names(res_mat) <- outcome

# Join with variable vector
plot_dat <- cbind.data.frame(vars, res_mat)

# Melt data (Convert to long form)
m_dat <- melt(plot_dat, id.vars = "vars")

# View melted data
#m_dat

g_1 <- ggplot(m_dat, aes(y = vars, x = variable, fill = value)) + # set aesthetics
  geom_tile() + # Use geom_tile for heatmap
  theme_bw() + # Set theme
  scale_fill_gradient2(low = "dark blue", # Choose low color
                       mid = "white", # Choose mid color
                       high = "red", # Choose high color
                       midpoint = mean(m_dat$value),
                       space = "Lab", 
                       na.value ="grey", # Choose NA value
                       guide = "colourbar", # Set color bar
                       aesthetics = "fill") + # Select aesthetics to apply
  labs(x = "Winner", y = "Characteristic", fill = "Mean value for winners and losers") # Set labels
g_1 # Generate plot

#cbind(res_mat, vars)
```
From the heat map it would appear that those who are winners are more likely to take down people, have significant strikes, and be on a current win streak than the ones who lost in the match.

```{r}
g_2 <- ggplot(combined[combined$weight_class %in% c("Heavyweight", "Featherweight", "Lightweight"),], aes(x = avg_SIG_STR_landed, fill = factor(Winner))) + # Set x as mean area and fill as diagnosis
  geom_density(alpha = 0.5) + # Select density plot and set transperancy (alpha)
    theme_set(theme_bw(base_size = 22) ) + # Set theme and text size
  facet_grid(~weight_class) +
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  labs(x = "Sign. Strike", title = "Significant Strikes - Winner v Loser",
       fill = "Winner") + # Set labels
  scale_fill_manual(values = c("1" = "red", "0" = "blue"), # Set fill colors manually
                    labels = c("1" = "Winner", "0" = "Loser")) # Set labels for fill
g_2 # Generate plot
```
From the g_2 density plot, it appears that losers have less significant strikes, while (surpisingly) in the lightweight category, the winners usually have less significant strikes.

```{r}
g_3<- ggplot(combined[combined$weight_class %in% c("Heavyweight", "Featherweight", "Lightweight"),], aes(x = avg_TD_att, fill = factor(Winner))) + # Set x as mean area and fill as diagnosis
  geom_density(alpha = 0.5) + # Select density plot and set transparency (alpha)
    theme_set(theme_bw(base_size = 22) ) + # Set theme and text size
  facet_grid(~weight_class) +
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  labs(x = "Average Takedown", title = "Takedowns - Winner v Loser",
       fill = "Winner") + # Set labels
  scale_fill_manual(values = c("1" = "red", "0" = "blue"), # Set fill colors manually
                    labels = c("1" = "Winner", "0" = "Loser")) # Set labels for fill
g_3 # Generate plot
```
It seems like that the losers and winners have approximately the same number of takedowns across weight classes, but it is important to mention that heavyweight fighters have less takedowns in general.

```{r}
g_4 <- ggplot(combined[combined$weight_class %in% c("Heavyweight", "Featherweight", "Lightweight"),], aes(x = win_by_KO.TKO, fill = factor(Winner))) + # Set x as mean area and fill as diagnosis
  geom_density(alpha = 0.5) + # Select density plot and set transparency (alpha)
    theme_set(theme_bw(base_size = 22) ) + # Set theme and text size
  facet_grid(~weight_class) +
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  labs(x = "WINS by TKO", title = "WINS by TKO - Winner v Loser",
       fill = "Winner") + # Set labels
  scale_fill_manual(values = c("1" = "red", "0" = "blue"), # Set fill colors manually
                    labels = c("1" = "Winner", "0" = "Loser")) # Set labels for fill
g_4 # Generate plot
```
There seems to be a correlation between prior TKO wins in heavyweight class and winners, unlike in the other weight classes.

## DEALING WITH AGE NAs
```{r}
combined <- drop_na(combined)
```

## New Variables
```{r}
combined$age_diff <- combined$age - combined$`age _opp`
combined$height_diff <- combined$Height_cms - combined$`Height_cms _opp`
combined$weight_diff <- combined$Weight_lbs - combined$`Weight_lbs _opp`
combined$reach_diff <- combined$Reach_cms - combined$`Reach_cms _opp`
```

## Formatting Variables
```{r}
## DATE TO DAYs
day_calc <- function(dates){
  x <- as.POSIXct(dates)
  return(floor(unclass(x)/86400))
}
combined$days <- day_calc(combined$date) # assign this to the dataset as days

## DATE TO MONTHs
combined$month <- month(combined$date) # assign this to the dataset as month

# Stance and Stance opponent as dummies 
combined <- dummy_cols(combined, select_columns = c("Stance", "Stance _opp"))
#combined

# Weightclass as dummies 
combined <- dummy_cols(combined, select_columns = c("weight_class"))
#combined


## Referee - Use top 5 and set the rest as "Other" for dummy variables
combined$Referee <- if_else(combined$Referee %in% c('Herb Dean', 'Dan Miragliotta', 'John McCarthy', 'Marc Goddard', 'Mario Yamasaki'), combined$Referee, "Other")
combined <- dummy_cols(combined, select_columns = c("Referee"))
titlebout <- rep(FALSE, nrow(combined))
titlebout[combined$title_bout == "True"] <- TRUE
combined$title_bout <- titlebout


## Deleting the unnecessary columns
combined <- combined[, !(colnames(combined) %in% c("location","date","Stance", "Stance _opp", "weight_class"))]
combined <-combined[, !(colnames(combined) %in% c("fighter","fighter _opp","draw","draw _opp", "Referee" ))]
names(combined)
```

```{r}
set.seed(1234)
split = sort(sample(nrow(combined), nrow(combined)*.8))
fights_train<-combined[split,]
fights_test<-combined[-split,]

# Cleaning up names for models
names(fights_train) <- make.names(names(fights_train))
names(fights_test) <- make.names(names(fights_test))

fights_train$Winner <- factor(fights_train$Winner)
fights_test$Winner <- factor(fights_test$Winner)
```

```{r}
# LASSO MODEL
set.seed(1234)
x_vars <- model.matrix(Winner ~., combined)[,-1]
combined_scaled <- scale(x_vars) # Scale dataset

# Fit lasso model with the best lambda
fit_1 <- glmnet(x = combined_scaled, # Set x variables
                y = combined$Winner, # Set response variable
                family = "binomial", # Set family to binomial
                alpha = 1, # Set alpha as 1 for lasso
                lambda = 0.01) # Set lambda as 0.01
# Print out coefficients
coef(fit_1)
```
___________________________________________________________________________________________________________________________________________________________________________
##Decision Tree
```{r}
# DECISION TREE
set.seed(1234) # Set random number generator seed for reproducibility
tree_model <- rpart(factor(Winner) ~., # Set tree formula
                data = fights_train) # Set dataset

# Prediction with classification tree
tree_preds <- predict(tree_model, fights_test, type = 'class')

#Confusion Matrix
t <- table(tree_preds, fights_test$Winner) # Create table
confusionMatrix(t)

# Accuracy: 0.5618
# Sensitivity: 0.5431
# Specificity: 0.5813
```
___________________________________________________________________________________________________________________________________________________________________________
##Bagging
```{r BootStrap Aggregration (Bagging)}
set.seed(1234) # Set random number generator seed for reproducibility

# Use random forest to do bagging
bag_mod <- randomForest(Winner ~., # Set tree formula
                data = fights_train, # Set dataset
                mtry = 164, # Set mtry to number of variables 
                ntree = 200) # Set number of trees to use
bag_mod # View model

# Confusion Matrix
bag_preds <- predict(bag_mod, fights_test) # Create predictions for bagging model

t <- table(bag_preds,fights_test$Winner) # Create table
confusionMatrix(t) # Produce confusion matrix
## ACCURACY: 0.5823 
## SENSITIVITY: 0.5847
```

```{r}
tree_1 <- getTree(bag_mod, 3, labelVar=TRUE) # Extract single tree,
head(tree_1) # Print first 5 rows of single tree details
```

```{r Error Rate}
oob_error <- bag_mod$err.rate[,1] # Extract oob error
plot_dat <- cbind.data.frame(rep(1:length(oob_error)), oob_error) # Create plot data
names(plot_dat) <- c("trees", "oob_error") # Name plot data


# Plot oob error
g_1 <- ggplot(plot_dat, aes(x = trees, y = oob_error)) + # Set x as trees and y as error
  geom_point(alpha = 0.5, color = "blue") + # Select geom point
  geom_smooth() + # Add smoothing line
  theme_bw() + # Set theme
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  labs(x = "Number of Trees", title = "Error Rate v Number of Trees",
       y = "Error Rate")  # Set labels
g_1 # Print plot

## number of trees = 180
```

```{r Bagging with 1000 trees}
set.seed(1234) # Set random number generator seed for reproducibility
bag_mod_2 <- randomForest(Winner ~., # Set tree formula
                data = fights_train, # Set data to use
                mtry = 164, # Set number of variables to use
                ntree = 1000) # Set number of trees

oob_error <- bag_mod_2$err.rate[,1] # Extract oob error
plot_dat <- cbind.data.frame(rep(1:length(oob_error)), oob_error) # Create plot data
names(plot_dat) <- c("trees", "oob_error") # Name plot data

# Plot oob error
g_2 <- ggplot(plot_dat, aes(x = trees, y = oob_error)) + # Set x as trees and y as error
  geom_point(alpha = 0.5, color = "blue") + # Select geom point
  geom_smooth() + # Add smoothing line
  theme_bw() + # Set theme
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  labs(x = "Number of Trees", title = "Error Rate v Number of Trees",
       y = "Error Rate")  # Set labels
g_2 # Create plot
## number of trees is arounf 800-ish
```

```{r 2nd Prediction Model Confusion Matrix}
bag_preds_2 <- predict(bag_mod_2, fights_test) # Predict test data

t <- table(bag_preds_2,fights_test$Winner) # Create table
confusionMatrix(t) # Produce confusion matrix
## ACCURACY: 0.5823 
## SENSITIVITY: 0.5833
```
### TAKES FOREVER
```{r Bagging Parameter Tuning}
#Could take a while to run 
trees <- c(10, 25, 50, 100, 200, 750) # Create vector of possible tree sizes
nodesize <- c(1, 10, 25, 50, 100, 200, 500, 1000) # Create vector of possible node sizes

params <- expand.grid(trees, nodesize) # Expand grid to get data frame of parameter combinations
names(params) <- c("trees", "nodesize") # Name parameter data frame
res_vec <- rep(NA, nrow(params)) # Create vector to store accuracy results

for(i in 1:nrow(params)){ # For each set of parameters
  set.seed(1234) # Set seed for reproducibility
  mod <- randomForest(Winner ~. , # Set formula
                      data=fights_train,# Set data
                      mtry = 164, # Set number of variables
                      importance = FALSE,  # 
                      ntree = params$trees[i], # Set number of trees
                      nodesize = params$nodesize[i]) # Set node size
  res_vec[i] <- 1 - mod$err.rate[nrow(mod$err.rate),1] # Calculate out of bag accuracy
}

summary(res_vec) # Summarize accuracy results

res_db <- cbind.data.frame(params, res_vec) # Join parameters and accuracy results
names(res_db)[3] <- "oob_accuracy" # Name accuracy results column
res_db # Print accuracy results column
```

```{r Heatmap Tuning}
res_db$trees <- as.factor(res_db$trees) # Convert tree number to factor for plotting
res_db$nodesize <- as.factor(res_db$nodesize) # Convert node size to factor for plotting
g_3 <- ggplot(res_db, aes(y = trees, x = nodesize, fill = oob_accuracy)) + # set aesthetics
  geom_tile() + # Use geom_tile for heatmap
  theme_bw() + # Set theme
  scale_fill_gradient2(low = "blue", # Choose low color
    mid = "white", # Choose mid color
    high = "red", # Choose high color
    midpoint =mean(res_db$oob_accuracy), # Choose mid point
    space = "Lab", 
    na.value ="grey", # Choose NA value
    guide = "colourbar", # Set color bar
    aesthetics = "fill") + # Select aesthetics to apply
  labs(x = "Node Size", y = "Number of Trees", fill = "OOB Accuracy") # Set labels
g_3 # Generate plot
```

```{r Best Results}
res_db[which.max(res_db$oob_accuracy),] # View best set of results
# TREES: 750
# NODE SIZE: 200
# OOB ACCURACY: 0.592717 
```

```{r Optimal Bagging}
set.seed(1234)
bag_mod_3 <- randomForest(Winner ~., # Set tree formula
                data = fights_train, # Set dataset
                mtry = 164, # Set number of variables 
                ntree = 750, # Set number of trees
                nodesize = 200) # Set node size

bag_preds_3 <- predict(bag_mod_3, fights_test) # Create predictions for test data


t <- table(bag_preds_3,  fights_test$Winner) # Create table
confusionMatrix(t) # Produce confusion matrix
```

```{r Cross Validation}
set.seed(1234) # Set seed for reproducibility
# Create cross-validation index
cv_ind <- sample(1:5, nrow(fights_train), replace = TRUE )

# Create accuracy store
cv_acc <- rep(NA, 5)
for(i in 1:5){ # For 1 to 5
  cv_train <- fights_train[cv_ind != i,] # Create training data
  cv_test <- fights_train[cv_ind == i,] # Create test data

  bag_mod_4 <- randomForest(Winner ~., # Set tree formula
                data = cv_train, # Set dataset
                mtry = 32, # set number of variables to use
                ntree = 500, # Set number of trees to generate
                nodesize = 10) # Set node size
  bag_preds_4 <- predict(bag_mod_4, cv_test) # Create test data predictions


  t <- table(bag_preds_4,cv_test$Winner) # Create table
  cf_mat <- confusionMatrix(t) # Create confusion matrix
  cv_acc[i] <- cf_mat$overall[1] # Extract accuracy
}

# Print cross validated accuracy scores
cv_acc

# Average Cross Validation
mean(cv_acc)
```

__________________________________________________________________________________________________________________________________________________________________________
##Random Forest

```{r Random Forest Training Model}
set.seed(1234)
head(fights_train)

rf_rate <- randomForest(Winner ~., # Set tree formula
                       data = fights_train, # Set dataset
                       ntree = 100) # Set number of trees to use
rf_rate # View model
```

```{r Random Forest Error Rate Plot}
oob_error1 <- rf_rate$err.rate[,1] # Extract oob error
plot_dat1 <- cbind.data.frame(rep(1:length(oob_error1)), oob_error1) # Create plot data
names(plot_dat1) <- c("trees", "oob_error") # Name plot data

# Plot oob error
rfoob <- ggplot(plot_dat1, aes(x = trees, y = oob_error)) + # Set x as trees and y as error
  geom_point(alpha = 0.5, color = "blue") + # Select geom point
  theme_bw() + # Set theme
  geom_smooth() + # Add smoothing line
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  labs(x = "Number of Trees", title = "Error Rate v Number of Trees",
       y = "Error Rate")  # Set labels
rfoob # Print plot
```

```{r Random Forest 1st Test Model}
rf_prediction <- predict(rf_rate, fights_test) # Create predictions for random forest model

rf_t <- table(rf_prediction, fights_test$Winner) # Create table
rf_t
confusionMatrix(rf_t) # Produce confusion matrix

# ACCURACY: 0.5859
# SENSITIVITY: 0.5667
```


```{r Random Forest Tuning}
rf_mtry_vals <- c(2, 4, 5, 7, 9, 12, 15, 20)
rf_nodesize_vals <- c(1, 10, 15, 50, 100, 150, 200, 500)

rf_params <- expand.grid(rf_mtry_vals, rf_nodesize_vals)
names(rf_params) <- c("mtry", "nodesize")
rf_acc_vec <- rep(NA, nrow(rf_params))
rf_sens_vec <- rep(NA, nrow(rf_params))

for(i in 1:nrow(rf_params)){
  set.seed(1234)
  rf_rate <- randomForest(Winner ~., # Set tree formula
                         data = fights_test, # Set dataset
                         ntree = 150,
                         nodesize = rf_params$nodesize[i],
                         mtry = rf_params$mtry[i]) # Set number of trees to use
  rf_preds_1 <-rf_rate$predicted # Create predictions for bagging model

  rf_t <- table(rf_preds_1,   fights_test$Winner) # Create table
  rf_c <- confusionMatrix(rf_t) # Produce confusion matrix
  
  rf_acc_vec[i] <- rf_c$overall[1]
  rf_sens_vec[i] <- rf_c$byClass[1]
}
```


```{r Random Forest Tuning Accuracy Visualization}
rf_res_db <- cbind.data.frame(rf_params, rf_acc_vec, rf_sens_vec)
rf_res_db$mtry <- as.factor(rf_res_db$mtry) # Convert tree number to factor for plotting
rf_res_db$nodesize <- as.factor(rf_res_db$nodesize) # Convert node size to factor for plotting
rf_g_1 <- ggplot(rf_res_db, aes(y = mtry, x = nodesize, fill = rf_acc_vec)) + # set aesthetics
  geom_tile() + # Use geom_tile for heatmap
  theme_bw() + # Set theme
  scale_fill_gradient2(low = "blue", # Choose low color
    mid = "white", # Choose mid color
    high = "red", # Choose high color
    midpoint =mean(rf_res_db$rf_acc_vec), # Choose mid point
    space = "Lab", 
    na.value ="grey", # Choose NA value
    guide = "colourbar", # Set color bar
    aesthetics = "fill") + # Select aesthetics to apply
  labs(x = "Node Size", y = "mtry", fill = "OOB Accuracy") # Set labels
rf_g_1 # Generate plot
```

```{r Random Forest Sensitivity Visualization}
rf_g_2 <- ggplot(rf_res_db, aes(y = mtry, x = nodesize, fill = rf_sens_vec)) + # set aesthetics
  geom_tile() + # Use geom_tile for heatmap
  theme_bw() + # Set theme
  scale_fill_gradient2(low = "blue", # Choose low color
    mid = "white", # Choose mid color
    high = "red", # Choose high color
    midpoint =mean(rf_res_db$rf_sens_vec), # Choose mid point
    space = "Lab", 
    na.value ="grey", # Choose NA value
    guide = "colourbar", # Set color bar
    aesthetics = "fill") + # Select aesthetics to apply
  labs(x = "Node Size", y = "Mtry", fill = "OOB Sensitivity") # Set labels
rf_g_2 # Generate plot
```

```{r Random Forest Determine Best Nodesize and Mtry}
rf_res_db[which(rf_res_db$nodesize == 1),]
## BEST:
# mtry: 15  nodesize: 1
```

```{r Finale Random Forest Model}
set.seed(1234)
rf_rate_final <- randomForest(Winner ~., # Set tree formula
                         data = fights_train, # Set dataset
                         ntree = 200,
                         nodesize = 1,
                         mtry = 15) # Set number of trees to use

rf_preds_1 <- predict(rf_rate_final,fights_test) # Create predictions for random forest model

rf_t <- table(rf_preds_1, fights_test$Winner) # Create table
confusionMatrix(rf_t) # Produce confusion matrix

## ACCURACY: 0.5739
## SENSITIVITY: 0.5625
## SPECIFICITY: 0.5856
```

___________________________________________________________________________________________________________________________________________________________________________
##XgBoost

```{r XGBoost Preparation}

set.seed(1234) # Set random number generator seed for reproducibility
#names(fights_train)
gtrain <- xgb.DMatrix(data = as.matrix(fights_train[,c(1:132,134:165)]), label = as.numeric(fights_train$Winner)-1)
# Create test matrix
gtest <- xgb.DMatrix(data = as.matrix(fights_test[,c(1:132,134:165)]), label = as.numeric(fights_test$Winner)-1)
```

```{r}
set.seed(1234)
fight_boost <- xgb.cv(data = gtrain, # Set training data
              
              nfold = 5, # Use 5 fold cross-validation
               
               eta = .3, # Set learning rate
              
               nrounds = 1000, # Set number of rounds
               early_stopping_rounds = 50, # Set number of rounds to stop at if there is no improvement
               
               verbose = 1, # 1 - Prints out fit
               nthread = 1, # Set number of parallel threads
               print_every_n = 20, # Prints out result every 20th iteration
              
               objective = "binary:logistic", # Set objective
               eval_metric = "auc",
               eval_metric = "error") # Set evaluation metric to use

gnb <- cbind.data.frame(fight_boost$evaluation_log[,c("iter", "test_error_mean")], rep(0.3, nrow(fight_boost$evaluation_log)))
names(gnb)[3] <- "eta"
```

```{r}
fight_1 <- ggplot(fight_boost$evaluation_log, aes(x = iter, y = test_error_mean)) + # Set x as trees and y as error
  geom_point(alpha = 0.5, color = "blue") + # Select geom point
  theme_bw() + # Set theme
  geom_smooth() + # Add smoothing line
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  labs(x = "Number of Trees", title = "Error Rate v Number of Trees",
       y = "Error Rate")  # Set labels
fight_1 # Print plot```
```

```{r}
set.seed(1234)
fight_boost1 <- xgboost(data = gtrain, # Set training data
              
              nfold = 5, # Use 5 fold cross-validation
               
               eta = 0.05, # Set learning rate
              
               nrounds = 200, # Set number of rounds
               early_stopping_rounds = 20, # Set number of rounds to stop at if there is no improvement
               
               verbose = 1, # 1 - Prints out fit
               nthread = 1, # Set number of parallel threads
               print_every_n = 20, # Prints out result every 20th iteration
              
               objective = "binary:logistic", # Set objective
               eval_metric = "auc",
               eval_metric = "error") # Set evaluation metric to use

```

```{r}
fight_preds <- predict(fight_boost1, gtest, type = "prob") # Create predictions

# Convert predictions to classes, using 0.5
fight_preds_class <- rep("0", length(fight_preds))
fight_preds_class[fight_preds >= 0.5] <- "1"


t <- table(fight_preds_class, fights_test$Winner) # Create table
confusionMatrix(t, positive = '1') # Produce confusion matrix
## ACCURACY: 0.5618
## SENSITIVITY: 0.7899
```

```{r}
imp_mat <- xgb.importance(model = fight_boost1)
#Plot importance (top 10 variables)
xgb.plot.importance(imp_mat, top_n = 10)

## MOST IMPORTANT VARIABLES:
# age_diff, avg_TD_att, avg_opp_SIG_STR_pct, avg_opp_SIG_STR_pct._opp, avg_SIG_STR_landed, avg_opp_TD_att, avg_TD_att._opp, avg_HEAD_landed, avg_opp_CTRL_time.seconds
```

```{r}
max_depth_vals <- c(3, 5, 7, 10, 15) # Create vector of max depth values
min_child_weight <- c(1,3,5,7, 10, 15) # Create vector of min child values

# Expand grid of parameter values
cv_params <- expand.grid(max_depth_vals, min_child_weight)
names(cv_params) <- c("max_depth", "min_child_weight")
# Create results vector
error_vec <- rep(NA, nrow(cv_params)) 


# Loop through results
for(i in 1:nrow(cv_params)){
  set.seed(1234)
  bst_tune <- xgb.cv(data = gtrain, # Set training data
              
              nfold = 5, # Use 5 fold cross-validation
              eta = 0.1, # Set learning rate
              max.depth = cv_params$max_depth[i], # Set max depth
              min_child_weight = cv_params$min_child_weight[i], # Set minimum number of samples in node to split
             
               
              nrounds = 100, # Set number of rounds
              early_stopping_rounds = 50, # Set number of rounds to stop at if there is no improvement
               
              verbose = 1, # 1 - Prints out fit
              nthread = 1, # Set number of parallel threads
              print_every_n = 20, # Prints out result every 20th iteration
               
              objective = "binary:logistic", # Set objective
              eval_metric = "error") # Set evaluation metric to use
  error_vec[i] <- bst_tune$evaluation_log$test_error_mean[bst_tune$best_ntreelimit]
}

```

```{r}
# Join results in dataset
res_db <- cbind.data.frame(cv_params, error_vec)
names(res_db)[3] <- "error" 
res_db$max_depth <- as.factor(res_db$max_depth) # Convert tree number to factor for plotting
res_db$min_child_weight <- as.factor(res_db$min_child_weight) # Convert node size to factor for plotting
# Print AUC heatmap
g_2 <- ggplot(res_db, aes(y = max_depth, x = min_child_weight, fill = error)) + # set aesthetics
  geom_tile() + # Use geom_tile for heatmap
  theme_bw() + # Set theme
  scale_fill_gradient2(low = "blue", # Choose low color
    mid = "white", # Choose mid color
    high = "red", # Choose high color
    midpoint =mean(res_db$error), # Choose mid point
    space = "Lab", 
    na.value ="grey", # Choose NA value
    guide = "colourbar", # Set color bar
    aesthetics = "fill") + # Select aesthetics to apply
  labs(x = "Minimum Child Weight", y = "Max Depth", fill = "Error") # Set labels
g_2 # Generate plot
```


## BASED ON THE BEST MIN CHILD WEIGHT (7) & MAX DEPTH (3), we will be trying to determine the eta:
```{r}

# Use xgb.cv to run cross-validation inside xgboost
set.seed(1234)
bst_mod_1 <- xgb.cv(data = gtrain, # Set training data
              
              nfold = 5, # Use 5 fold cross-validation
               
              eta = 0.3, # Set learning rate
              max.depth = 3, # Set max depth
              min_child_weight = 7, # Set minimum number of samples in node to split
              gamma = 0, # Set minimum loss reduction for split
              subsample = 0.9, # Set proportion of training data to use in tree
              colsample_bytree =  0.9, # Set number of variables to use in each tree
               
              nrounds = 1000, # Set number of rounds
              early_stopping_rounds = 20, # Set number of rounds to stop at if there is no improvement
               
              verbose = 1, # 1 - Prints out fit
              nthread = 1, # Set number of parallel threads
              print_every_n = 20, # Prints out result every 20th iteration
              
              objective = "binary:logistic", # Set objective
              eval_metric = "auc",
              eval_metric = "error") # Set evaluation metric to use


set.seed(1234)
bst_mod_2 <- xgb.cv(data = gtrain, # Set training data
              
              nfold = 5, # Use 5 fold cross-validation
               
              eta = 0.1, # Set learning rate
              max.depth =  3, # Set max depth
              min_child_weight = 7, # Set minimum number of samples in node to split
              gamma = 0, # Set minimum loss reduction for split
              subsample = 0.9 , # Set proportion of training data to use in tree
              colsample_bytree = 0.9, # Set number of variables to use in each tree
               
              nrounds = 1000, # Set number of rounds
              early_stopping_rounds = 20, # Set number of rounds to stop at if there is no improvement
               
              verbose = 1, # 1 - Prints out fit
              nthread = 1, # Set number of parallel threads
              print_every_n = 20, # Prints out result every 20th iteration
              
              objective = "binary:logistic", # Set objective
              eval_metric = "auc",
              eval_metric = "error") # Set evaluation metric to use
set.seed(1234)
bst_mod_3 <- xgb.cv(data = gtrain, # Set training data
              
              nfold = 5, # Use 5 fold cross-validation
               
              eta = 0.05, # Set learning rate
              max.depth = 3, # Set max depth
              min_child_weight = 7 , # Set minimum number of samples in node to split
              gamma = 0, # Set minimum loss reduction for split
              subsample = 0.9 , # Set proportion of training data to use in tree
              colsample_bytree =  0.9, # Set number of variables to use in each tree
               
              nrounds = 1000, # Set number of rounds
              early_stopping_rounds = 20, # Set number of rounds to stop at if there is no improvement
               
              verbose = 1, # 1 - Prints out fit
              nthread = 1, # Set number of parallel threads
              print_every_n = 20, # Prints out result every 20th iteration
              
              objective = "binary:logistic", # Set objective
              eval_metric = "auc",
              eval_metric = "error") # Set evaluation metric to use
set.seed(1234)
bst_mod_4 <- xgb.cv(data = gtrain, # Set training data
              
              nfold = 5, # Use 5 fold cross-validation
               
              eta = 0.01, # Set learning rate
              max.depth = 3, # Set max depth
              min_child_weight = 7, # Set minimum number of samples in node to split
              gamma = 0.1, # Set minimum loss reduction for split
              subsample = 0.9 , # Set proportion of training data to use in tree
              colsample_bytree = 0.9, # Set number of variables to use in each tree
               
              nrounds = 1000, # Set number of rounds
              early_stopping_rounds = 20, # Set number of rounds to stop at if there is no improvement
               
              verbose = 1, # 1 - Prints out fit
              nthread = 1, # Set number of parallel threads
              print_every_n = 20, # Prints out result every 20th iteration
              
              objective = "binary:logistic", # Set objective
              eval_metric = "auc",
              eval_metric = "error") # Set evaluation metric to use

set.seed(1234)
bst_mod_5 <- xgb.cv(data = gtrain, # Set training data
              
              nfold = 5, # Use 5 fold cross-validation
               
              eta = 0.005, # Set learning rate
              max.depth = 3, # Set max depth
              min_child_weight = 7, # Set minimum number of samples in node to split
              gamma = 0, # Set minimum loss reduction for split
              subsample = 0.9 , # Set proportion of training data to use in tree
              colsample_bytree = 0.9, # Set number of variables to use in each tree
               
              nrounds = 1000, # Set number of rounds
              early_stopping_rounds = 20, # Set number of rounds to stop at if there is no improvement
               
              verbose = 1, # 1 - Prints out fit
              nthread = 1, # Set number of parallel threads
              print_every_n = 20, # Prints out result every 20th iteration
               
              objective = "binary:logistic", # Set objective
              eval_metric = "auc",
              eval_metric = "error") # Set evaluation metric to use
```

```{r}
pd1 <- cbind.data.frame(bst_mod_1$evaluation_log[,c("iter", "test_error_mean")], rep(0.3, nrow(bst_mod_1$evaluation_log)))
names(pd1)[3] <- "eta"
# Extract results for model with eta = 0.1
pd2 <- cbind.data.frame(bst_mod_2$evaluation_log[,c("iter", "test_error_mean")], rep(0.1, nrow(bst_mod_2$evaluation_log)))
names(pd2)[3] <- "eta"
# Extract results for model with eta = 0.05
pd3 <- cbind.data.frame(bst_mod_3$evaluation_log[,c("iter", "test_error_mean")], rep(0.05, nrow(bst_mod_3$evaluation_log)))
names(pd3)[3] <- "eta"
# Extract results for model with eta = 0.01
pd4 <- cbind.data.frame(bst_mod_4$evaluation_log[,c("iter", "test_error_mean")], rep(0.01, nrow(bst_mod_4$evaluation_log)))
names(pd4)[3] <- "eta"
# Extract results for model with eta = 0.005
pd5 <- cbind.data.frame(bst_mod_5$evaluation_log[,c("iter", "test_error_mean")], rep(0.005, nrow(bst_mod_5$evaluation_log)))
names(pd5)[3] <- "eta"
# Join datasets
plot_data <- rbind.data.frame(pd1, pd2, pd3, pd4, pd5)
# Converty ETA to factor
plot_data$eta <- as.factor(plot_data$eta)
# Plot points
g_6 <- ggplot(plot_data, aes(x = iter, y = test_error_mean, color = eta))+
  geom_point(alpha = 0.5) +
  theme_bw() + # Set theme
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  labs(x = "Number of Trees", title = "Error Rate v Number of Trees",
       y = "Error Rate", color = "Learning \n Rate")  # Set labels
g_6
```


## WE WILL END UP WITH THE BEST ETA (.1), number of rounds, max depth of 15, and min child weight of 1.

```{r}
set.seed(1234)
bst_final <- xgboost(data = gtrain, # Set training data
              
        
               
              eta = 0.1, # Set learning rate
              max.depth =  3, # Set max depth
              min_child_weight = 7, # Set minimum number of samples in node to split
              gamma = 0, # Set minimum loss reduction for split
              subsample =  0.9, # Set proportion of training data to use in tree
              colsample_bytree = 0.9, # Set number of variables to use in each tree
               
              nrounds = 100, # Set number of rounds
              early_stopping_rounds = 20, # Set number of rounds to stop at if there is no improvement
               
              verbose = 1, # 1 - Prints out fit
              nthread = 1, # Set number of parallel threads
              print_every_n = 20, # Prints out result every 20th iteration
              
              objective = "binary:logistic", # Set objective
              eval_metric = "auc",
              eval_metric = "error") # Set evaluation metric to use


fight_preds_best <- predict(bst_final, gtest, type = "prob") # Create predictions

# Convert predictions to classes, using 0.45
fight_preds_best_class <- rep("0", length(fight_preds_best))
fight_preds_best_class[fight_preds_best >= 0.45] <- "1"


t <- table(fight_preds_best_class, fights_test$Winner) # Create table
confusionMatrix(t, positive = '1')
# CUTOFF - 0.45
# ACCURACY: 0.6035
# SENSITIVITY: 0.7554
# SPECIFICITY: 0.4569

# CUTOFF - 0.55
# ACCURACY: 0.5788
# SENSITIVITY: 0.4388
# SPECIFICITY: 0.7139
```

## VARIABLE IMPORTANCE PLOTS BASED ON OUR Best MODEL
```{r}
imp_mat <- xgb.importance(model = bst_final)
#Plot importance (top 10 variables)
xgb.plot.importance(imp_mat, top_n = 10)

## 10 MOST IMPORTANT VARIABLES:
# age_diff, reach_diff, longest_win_streak, avg_SIG_STR_landed_opp, avg_HEAD_landed_opp, longest_win_streak_opp, avg_TD_att, avg_opp_SIG_STR_pct, avg_SIG_STR_landed, avg_HEAD_landed
```

____________________________________________________________________________________________________________________________________________________________________________
## SHAP PLOT

```{r}
source("a_insights_shap_functions.r")
```

```{r}
shap_result_1 <- shap.score.rank(xgb_model = bst_final, 
                X_train = as.matrix(fights_train[,c(1:132,134:165)]),
                shap_approx = F)
```

```{r}
shap_long_1 = shap.prep(shap = shap_result_1,
                           X_train = as.matrix(fights_train[,c(1:132,134:165)]), 
                           top_n = 10)
```

```{r}
plot.shap.summary(data_long = shap_long_1)
```

```{r}
# Exclude Non-Controllable Variables from G-Train and G-Test
fights_train_exclude <- fights_train[, !(colnames(fights_train) %in% c("total_title_bouts", "total_time_fought.seconds.", "total_rounds_fought", "total_title_bouts", "current_win_streak", "current_lose_streak", "longest_win_streak", "wins", "losses", "win_by_Decision_Majority", "win_by_Decision_Split", "win_by_Decision_Unanimous", "win_by_KO.TKO", "win_by_Submission", "win_by_TKO_Doctor_Stoppage", "Height_cms", "Reach_cms", "age", "total_time_fought.seconds.._opp", "total_rounds_fought._opp", "total_title_bouts._opp", "current_win_streak._opp", "current_lose_streak._opp", "longest_win_streak._opp", "wins._opp", "losses._opp", "win_by_Decision_Majority._opp", "win_by_Decision_Split._opp", "win_by_Decision_Unanimous._opp", "win_by_KO.TKO._opp", "win_by_Submission._opp", "win_by_TKO_Doctor_Stoppage._opp", "Height_cms._opp", "Reach_cms._opp", "age._opp", "title_bout", "age_diff", "height_diff", "weight_diff", "reach_diff", "days", "month", "Referee_Dan.Miragliotta", "Referee_Herb.Dean", "Referee_John.McCarthy", "Referee_Marc.Goddard", "Referee_Mario.Yamasaki", "Referee_Other"))]
fights_test_exclude <- fights_test[, !(colnames(fights_test) %in% c("total_title_bouts", "total_time_fought.seconds.", "total_rounds_fought", "total_title_bouts", "current_win_streak", "current_lose_streak", "longest_win_streak", "wins", "losses", "win_by_Decision_Majority", "win_by_Decision_Split", "win_by_Decision_Unanimous", "win_by_KO.TKO", "win_by_Submission", "win_by_TKO_Doctor_Stoppage", "Height_cms", "Reach_cms", "age", "total_time_fought.seconds.._opp", "total_rounds_fought._opp", "total_title_bouts._opp", "current_win_streak._opp", "current_lose_streak._opp", "longest_win_streak._opp", "wins._opp", "losses._opp", "win_by_Decision_Majority._opp", "win_by_Decision_Split._opp", "win_by_Decision_Unanimous._opp", "win_by_KO.TKO._opp", "win_by_Submission._opp", "win_by_TKO_Doctor_Stoppage._opp", "Height_cms._opp", "Reach_cms._opp", "age._opp", "title_bout", "age_diff", "height_diff", "weight_diff", "reach_diff", "days", "month", "Referee_Dan.Miragliotta", "Referee_Herb.Dean", "Referee_John.McCarthy", "Referee_Marc.Goddard", "Referee_Mario.Yamasaki", "Referee_Other"))]

set.seed(1234) # Set random number generator seed for reproducibility
#names(fights_train)
gtrain_exclude <- xgb.DMatrix(data = as.matrix(fights_train_exclude[,c(1:98,100:118)]), label = as.numeric(fights_train_exclude$Winner)-1)
# Create test matrix
gtest_exclude <- xgb.DMatrix(data = as.matrix(fights_test_exclude[,c(1:98,100:118)]), label = as.numeric(fights_test_exclude$Winner)-1)

colnames(fights_train_exclude)
```

```{r}
# Run New XgBoost Model
set.seed(1234)
bst_final_2 <- xgboost(data = gtrain_exclude, # Set training data
              
        
               
              eta = 0.1, # Set learning rate
              max.depth =  3, # Set max depth
              min_child_weight = 7, # Set minimum number of samples in node to split
              gamma = 0, # Set minimum loss reduction for split
              subsample =  0.9, # Set proportion of training data to use in tree
              colsample_bytree = 0.9, # Set number of variables to use in each tree
               
              nrounds = 100, # Set number of rounds
              early_stopping_rounds = 20, # Set number of rounds to stop at if there is no improvement
               
              verbose = 1, # 1 - Prints out fit
              nthread = 1, # Set number of parallel threads
              print_every_n = 20, # Prints out result every 20th iteration
              
              objective = "binary:logistic", # Set objective
              eval_metric = "auc",
              eval_metric = "error") # Set evaluation metric to use


fight_preds_2 <- predict(bst_final_2, gtest_exclude, type = "prob") # Create predictions

# Convert predictions to classes, using 0.55
fight_preds_2_class <- rep("0", length(fight_preds_2))
fight_preds_2_class[fight_preds_2 >= 0.55] <- "1"


t_2 <- table(fight_preds_2_class, fights_test_exclude$Winner) # Create table
confusionMatrix(t_2, positive = '1')
```

```{r}
shap_result_2 <- shap.score.rank(xgb_model = bst_final_2, 
                X_train = as.matrix(fights_train_exclude[,c(1:98,100:118)]),
                shap_approx = F)
```

```{r}
shap_long_2 = shap.prep(shap = shap_result_2,
                           X_train = as.matrix(fights_train_exclude[,c(1:98,100:118)]), 
                           top_n = 10)
```

```{r}
plot.shap.summary(data_long = shap_long_2)
```


```{r}
# Calculate initial model ROC
roc1 = roc(fights_test$Winner, fight_preds)
# Calculate performance vars model ROC
roc2 = roc(fights_test$Winner, fight_preds_2)
# Calculate performance vars model ROC
roc3 = roc(fights_test$Winner, fight_preds_best)
# Print initial model AUC
plot.roc(roc1, print.auc = TRUE, print.auc.x = 0.25, print.auc.y = 0.6, col = "red", print.auc.col = "red")
# Print performance vars model AUC
plot.roc(roc2, print.auc = TRUE, print.auc.x = 0.25, print.auc.y = 0.4, col ="blue", print.auc.col = "blue", add = TRUE)
# Print best model AUC
plot.roc(roc3, print.auc = TRUE, print.auc.x = 0.25, print.auc.y = 0.2, col ="green", print.auc.col = "green", add = TRUE)
```

```{r}
# Prepare for ROC/AUC
tree_preds <- predict(tree_model, fights_test, type = 'prob')
rf_preds_1 <- predict(rf_rate, fights_test, type = 'prob')

# Calculate decision tree model ROC
roc1 = roc(fights_test$Winner, tree_preds[,2])
# Calculate random forest model ROC
roc2 = roc(fights_test$Winner, rf_preds_1[,2])
# Calculate XgBoost model ROC
roc3 = roc(fights_test$Winner, fight_preds_best)
# Print decision tree model AUC
plot.roc(roc1, print.auc = TRUE, print.auc.x = 0.25, print.auc.y = 0.6, col = "red", print.auc.col = "red")
# Print random forest model AUC
plot.roc(roc2, print.auc = TRUE, print.auc.x = 0.25, print.auc.y = 0.4, col ="blue", print.auc.col = "blue", add = TRUE)
# Print XgBoost model AUC
plot.roc(roc3, print.auc = TRUE, print.auc.x = 0.25, print.auc.y = 0.2, col ="green", print.auc.col = "green", add = TRUE)
```